### 2019-06-11 report/chapter01.Rnw
Methods update

>>>>

Typically, we expect empirical scientific results to be distorted by random noise, such that the results are not equal to the real process that they describe. Additionally, one can reasonably assume that the results are also biased to some extent, as the statistical paradigm of precision versus bias would predict: The smaller the noise part of distortion, the more the result should be distorted by bias. However, the trade-off is not inevitable and can be minimized by scientific rigour and foresightful experimental design. Large efforts have been made both in theory and practice to improve the quality of experiments and their analysis, such as defining experimental settings that prohibit influence of expectations of the researcher (e.g. blinding) or increase sample size in experiments. The introduction of randomized clinical trials is for example seen nowadays as a benchmark in clinical science, heavily improving the reliability of its findings. \\
It is more and more considered a new scientific field by itself to think of and argue about circumstances that improve the quality of scientific findings. Statistics is in some sense predestined to contribute to it, because probability and chance concepts provide are already at the core of understanding of each empirical science. \\
There are however issues in which statistics is not thought to be able to contribute much understanding and to offer little help. For example, the selective attention that agents in modern science are paying to new, strong and clear findings in each science is considered merely a psychological, socioeconomic issue. It is often explained by the affinity of humans for stories rather than mere facts and for novelties. This issue, that scientific findings get more attention, and today, are more likely to be published, read and cited, is known for many years now, but has only gained new traction in meta science when the reproducibility crisis emerged in 2003 to 2005. It was the studies of ... that showed that even rigid prosecution of study protocolls did often not lead to reproduction of previous results when experiments were repeated, and struck empirical science in its core. Most reasonably, ... concluded that it was a large missunderstanding of statistical tools such as the p-value in the scientific community that lead to this situation. Although that some measures have been taken since then, and some progress is made, the non-reproducibility of scientific fiendings remain an accute threat to the relevance and reliability of science. Some even argue that certain scientific fields sooner or later will repeat the experience that for example psychology or medicine have made, and are yet to encounter their own reproducibility crisis. \\
Clearly, reproduction of experimental results is the gold standard of testing empirical science, because it reassures the universality and reliability by which certain procedures or effects appear if measured in the correct experimental context. However, there are also problems there. One can almost never fully exclude chance events to have played a major role in negating or reafirming the experimental findings (which is of course always the case where noise is present in the data). Furthermore, reproduction studies are generally costly and the precise experimental conditions might be hard to reproduce because of lack of information or other reasons. There is another way to assess the strength of scientific results if the experiment has been conducted more than once (however not with the same study protocoll as in reproduction studies): Meta-analyses. In a meta-analyses, results of multiple (fairly identical) experiments (studies) are summarized to a single result. Very often, meta-analyses are not only a synthesis of results, but of evidence, thus reflecting the overall and summarized evidence regarding to a scientific question. It is on purpose that selective attention to positive, strong results in the scientific literature have been mentioned beforehand, because it is now clear how they will affect this second way of reassurance of scientific findigs: The meta-analysis will again lead to irreproducible findings if it is based on a set of results that are all positive since they are easier available and more prominent in the scientific literature. Thus, meta-analyses will in this case not reach their purpose of assessing reliability of science, but worsen the problem by reinforcing the confidence in the overestimated effects in the experiment. However, and this is the main topic in this masters thesis, there are some ways to detect irregularities in the body of results that can provide indirect hints that a selective rather than a random sample out of a hypothetical population of experiments is present. The abundance of the methods to link some features of the sample of experiments to publication bias, as the tendency of scientific literature to include over-proportionally large effects is generally termed, also speaks for the relevance of the topic. A review in 2017 identified alone 147 (!) conceptually differing methods.\\
This speaks as well for the inevitable difficulties that such methods encounter. First of all, it is most often impossible to estimate the real selection process, that is the rate by which smaller effects go unnoticed because of selective publication and reporting, because the number of missing results in the studies is not known at the first place and impossible to retrieve. Secondly, there is almost no real world data of complete and unbiased meta-analyses, such that evaluations of methods is most often dependent on simulations. So we have, after applying the methods, only indirect information of publication bias, and the extent to what it might influences scientific findings. There are, almost all the time, alternative explanations for the results of the methods that relieve the operators and publishers from the reproval of publication bias. However, given the large body of evidence for publication bias to play a role in science that has been collected in other ways, those can be expelled most of the time, such that those methods can give a quantitative measure of publication bias. Before beginning with the main part of this masters thesis, I want to recall some of the most pointed evidence for publication bias that has been collected for publication bias so far. \\
